---
title: "Covid91 vaccine study Final2023"
author: "Jonathan Urbano"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, covid19 = "covid", OK = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
FinalData$treatment <- relevel(FinalData$treatment,ref="placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")
Males = filter(FinalData, sex =="M")
Females = filter(FinalData, sex =="F")
LGBTQ = filter(FinalData, LGBTQ =="gay")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## Barplot 
```{r}
dd2 <- Males %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This barplot shows the number of men who got Covid after taking the vaccine and placebo. The boxplot above shows that 30% of people who received the vaccine got COVID-19, compared to 69% of people who received the placebo. A significantly higher percentage of those who received the placebo developed covid-19 compared to those who received the vaccine. The vaccine appears to be effective in protecting against Covid-19 disease. The odds of not getting covid-19 while on the vaccine are 2.25:1. Men who received the vaccine were 2.25 times less likely to get COVID-19 than those who received a placebo. In the chart, the two bars on the left side of the graph show all the people who have been infected with covid-19. The X-axis shows whether the person is infected with covid-19 or not. However, the y-axis does not show the number of people infected with COVID-19, but now the percentage of people who got covid-19 after taking the placebo or the vaccine. The codes that generate this diagram begin with a block of code beginning with dd2. dd2 appears in the environment window and when checked contains 4 variables. These variables include infected, treatment, number and percentage (represents a percentage). The second code group is called baseC. BasicC, the last set of code, can also be seen in the environment window and appears to contain a list of 9 variables. This information creates the first plot. 
  
### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=Males)
rowPerc(table1)
colPerc(table1)
```
According to the numerical results, 69% of the men got sick after taking the placebo and only 30% of the men after the vaccine. A significantly higher percentage of patients who received the placebo contracted covid-19 than those who received the vaccine. The vaccine appears to offer better protection against Covid-19. There is a 2.25:1 chance that you will not get the coronavirus while you get the vaccine. Men who received the vaccine were 2.25 times less likely to develop COVID-19 than men who received a placebo..
### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
Malesfish <- fisher.test(table1)
Malesfish
```
Among all these results, the result of Fisher's exact test must be considered. The p-value above is very small at 1.098e-13. This is a very small number. seeing this number means that these results are unlikely to come by chance. The null hypothesis predicted that deaths related to covid-19 would be independent of gender. Since it is clear that gender plays a role in individual mortality, the alternative hypothesis that there is a correlation between COVID-19 deaths and gender is a favorable choice. The odd ratio above describes the ratio 2.242522. men (men) who received the vaccine were 2.24 times less likely to get COVID-19 than those who received the placebo.
```{r}
effect <- Malesfish$p.value/(Malesfish$p.value +1)
effectivenessMales = Malesfish$estimate/(Malesfish$estimate +1)*100
print(effectivenessMales)
```
```{r}
low=(Malesfish$conf.int[1]/(Malesfish$conf.int[1] +1))*100
hi=(Malesfish$conf.int[2]/(Malesfish$conf.int[2] +1))*100
options(digits=2)
```
## Effectiveness Confidence Interval `r low`  to `r hi`
Above is the effectiveness of the male (male) covid-19 vaccine. If the strength is above 50%, it means that the vaccine is effective. Based on the above period, we can conclude that the vaccine effectively protects men against covid-19 infection.
The probability shows how effective the vaccine is. The odds ratio for men is 69%.




## Females
```{r}
dd2 <- Females %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This graph shows the number of women who got Covid after taking the vaccine and a placebo. The plot shows that more people who received only the placebo contracted covid-19. Only about 40 percent of people who received the vaccine were infected with covid-19. This means that about 60% of those infected with COVID-19 were those who received only the placebo. The probability of contracting COVID-19 while receiving the vaccine is 1.5. Women who received the vaccine were 1.5 times less likely to develop covid-19 than women who received a placebo. The two bars on the left of the graph show all the people who have been infected with covid-19. The X-axis shows whether the person is infected with covid-19 or not. However, the y-axis does not show the number of people infected with COVID-19, but now the percentage of people who got covid-19 after taking the placebo or the vaccine. The codes that generate this diagram begin with a block of code that begins with dd2. dd2 appears in the environment window and on inspection it contains 4 variables. These variables are infected, treatment, count and percentage (means percentage). The second code group is called baseC. BasicC can also be seen in the environment window and appears to contain a list of 9 variables. This information creates the first plot. BasicCC is the last set of code and is also in the file window..

### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```
According to the graph and numerical results, 60% of the women got sick after taking the placebo and only 40% of the women after the vaccine. A significantly higher percentage of patients who received the placebo contracted covid-19 than those who received the vaccine. The vaccine appears to offer better protection against Covid-19. Women have a 1.5 chance of getting COVID-19 while receiving the vaccine. Women who received the vaccine were 1.5 times less likely to develop covid-19 than women who received a placebo..
### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
Femalesfish <- fisher.test(table1)
Femalesfish
```
Among all these results, the result of Fisher's exact test must be considered. The p-value shown above is a very small 3e-04. This is a very small number. seeing this number means that these results are unlikely to occur by chance, so the null hypothesis can be rejected in favor of the alternative. Fisher's exact test shows that the probability is 1.5. Women who received the vaccine were about 1.5 times less likely to get COVID-19 than women who received only the placebo..
```{r}
effect <- Femalesfish$p.value/(Femalesfish$p.value +1)
effectivenessFemales = Femalesfish$estimate/(Femalesfish$estimate +1)*100
print(effectivenessFemales)
```
```{r}
low=(Femalesfish$conf.int[1]/(Femalesfish$conf.int[1] +1))*100
hi=(Femalesfish$conf.int[2]/(Femalesfish$conf.int[2] +1))*100
options(digits=2)
```
## Effectiveness Confidence Interval `r low`  to `r hi`
Above is the effectiveness of the COVID-19 vaccine for women. If the strength is above 50%, it means that the vaccine is effective. Based on the above period, we can conclude that the vaccine effectively protects Femaales against covid-19 infection. The probability is approximately the effectiveness of the vaccine. For women, the odds ratio is 60%.






## LGBTQ
```{r}
dd2 <- LGBTQ %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))
# the group_by followed by summarize(count=n())
basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")
#Now for percentage plot
basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This graph shows the number of LGBT people who got Covid after taking the vaccine and a placebo. This line chart shows that 85% of the people who got the vaccine got covid-19 and 15% of the people who got the placebo. The probability of contracting COVID-19 while receiving the vaccine is 0.18. People who identified as gay and received the vaccine were 0.18 times more likely to get covid-19 than those who received a placebo. There are two plots. In the first graph, the x-axis represents those infected, regardless of whether the participants were infected with covid-19 or not. The y-axis represents the number of people infected with COVID-19. In the second graph above, the two bars on the left of the graph are all the people who have been infected with covid-19. The X-axis shows whether the person is infected with covid-19 or not. However, the y-axis does not show the number of people infected with COVID-19, but now the percentage of people who got covid-19 after taking the placebo or the vaccine. The codes that generate this diagram begin with a block of code beginning with dd2. dd2 appears in the environment window and on inspection it contains 4 variables. These variables are infected, treatment, count and percentage (means percentage). The second code group is called baseC. BasicC can also be seen in the environment window and appears to contain a list of 9 variables. This information creates the first plot. BasicCC is the last set of code and is also in the file window
### Numerical results
```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
These numerical results show the specific percentage of LGBT people who got COVID-19 after taking the placebo and the vaccine. The percentage of LGBTQ participants who took the vaccine and contracted covid-19 was a shocking 85 percent! Only 15 percent of the participants who received only the placebo contracted covid-19. Vaccine ID is apparently completely ineffective in protecting the LGBTQ community from covid-19 and is not recommended. The probability of contracting COVID-19 while receiving the vaccine is 0.18. People who identified as LGBTQ and received the vaccine were 0.18 times more likely to get Covid-19 than people who received a placebo.
### Inferential 
```{r}
chisq.test(table1)
chisqtestGC(table1)
LGBTQfish <- fisher.test(table1)
LGBTQfish
```
The p-value, standing at 7e-08, represents an exceedingly minute probability. Consequently, the inference drawn is that these outcomes are highly improbable to have arisen randomly. The Fishers's Exact Test reveals an odds ratio of 0.18, indicating that LGBTQ individuals vaccinated were approximately 0.18 times as likely to contract Covid-19 compared to those who received the placebo. Rejecting the null hypothesis in favor of the alternative, the drug demonstrates complete ineffectiveness in mitigating Covid-19 risks among individuals identifying as gay.
```{r}
effect <- LGBTQfish$p.value/(LGBTQfish$p.value +1)
effectivenessLGBTQ = LGBTQfish$estimate/(LGBTQfish$estimate +1)*100
print(effectivenessLGBTQ)
```
```{r}
low=(LGBTQfish$conf.int[1]/(LGBTQfish$conf.int[1] +1))*100
hi=(LGBTQfish$conf.int[2]/(LGBTQfish$conf.int[2] +1))*100
options(digits=2)
```
## Effectiveness Confidence Interval `r low`  to `r hi`
The efficacy of the Covid-19 vaccine among individuals identifying as LGBTQ is presented above. When effectiveness surpasses 50%, it signifies the vaccine's efficacy. However, based on the interval provided, it's evident that the vaccine does not offer protection against Covid-19 for LGBTQ individuals. The odds ratio serves as a gauge of the vaccine's effectiveness, which for the LGBTQ population is approximately 15, indicating complete ineffectiveness. Those who receive the vaccine aren't at an advantage compared to those who don't receive it. The confidence interval for effectiveness among LGBTQ individuals spans from 7.6 to 26.57.


# Overall Results and Conclusions
The Covid-19 vaccine showed ineffectiveness for both LGBTQ individuals and individuals with a history of drug use, as their effectiveness percentages fell below 50%. Conversely, the vaccine proved effective for both males and females, with effectiveness scores surpassing 50%. For males, the Covid-19 vaccine demonstrated efficacy in preventing the contraction of the virus, with an effectiveness rate of r effectivenessMales. Men benefited the most from this particular vaccine. Similarly, the Covid-19 vaccine exhibited high success rates in protecting women against the virus, with an effectiveness score of r effectivenessFemales. However, the vaccine was not effective for individuals who identified as LGBTQ. Those who received the vaccine were 0.18 times more likely to contract Covid-19 compared to those taking the placebo. The effectiveness rate for LGBTQ individuals was r effectivenessLGBTQ. Likewise, individuals with a history of drug use who received the vaccine were 0.13 times more likely to contract Covid-19 than those taking the placebo. The effectiveness rate for this group was r effectivenessDruggies. It is not recommended for individuals who identify as LGBTQ or have a history of drug use to receive the vaccine. Conversely, for males and females, the vaccine was reported to be effective. The null hypothesis, suggesting no difference in vaccine efficacy, can be accepted for categories including individuals with a history of drug use and LGBTQ individuals. However, the alternative hypothesis, indicating the vaccine's efficacy, can be accepted for males and females.




